---
title: "Reproducible Research - Course Project 1"
author: "Yoann Pannetier"
date: "7/1/2020"
output: html_document
---

Loading required libraries
==============

```{r}
library(knitr)
library(ggplot2)
```

1. Code for reading in the dataset and/or processing the data
==============

Load the CSV file into a R variable (data)

```{r}
data <- read.csv('activity.csv')
```

Transform the date variable in data into a R date class (format YYY-MM-DD)

```{r}
data <- transform(data, date = as.Date(date))
```

2. Histogram of the total number of steps taken each day
==============

Aggregate the sum of steps for each day and remove invalid numbers (NA)

```{r}
steps_per_day <- aggregate(steps ~ date, data, sum, na.rm = TRUE)
```

Create histogram plot of the numer of days for each range of numbers of steps

```{r}
hist(steps_per_day$steps, main = "Total Steps per day", xlab = "Steps", ylim = c(0,40), labels = TRUE)
```

3. Mean and median number of steps taken each day
==============

Calculate the mean and median of total steps taken per day: 

- mean:

```{r}
mean(steps_per_day$steps)
```

- median:

```{r}
median(steps_per_day$steps)
```

4. Time series plot of the average number of steps taken
==============

Aggregate the mean of steps for each day and remove invalid numbers (NA)

```{r}
mean_steps_interval <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
```

Create time series plot

```{r}
plot(mean_steps_interval$interval, mean_steps_interval$steps, type = "l", main = "Average number of steps per 5-min interval", xlab = "Interval", ylab = "Average Steps")
```

5. The 5-minute interval that, on average, contains the maximum number of steps
==============

Select the 5-minute interval with the maximum value for steps

```{r}
max_interval <- mean_steps_interval$interval[which(mean_steps_interval$steps == max(mean_steps_interval$steps))]
print(max_interval)
```

6. Code to describe and show a strategy for imputing missing data
==============

Report number of missing values in the dataset

```{r}
sum(is.na(data))
```

Replace NA with average value of the 5-min interval

```{r}
imputed_data <- data
for (i in 1:length(imputed_data$steps)) {
  if (is.na(imputed_data$steps[i])) {
    imputed_data$steps[i] <- mean_steps_interval$steps[mean_steps_interval$interval == imputed_data$interval[i]]
  }
}
```

Aggregate the data et report number of missing values in the dataset as a control

```{r}
imp_steps_per_day <- aggregate(steps ~ date, imputed_data, sum, na.rm = TRUE)
sum(is.na(imp_steps_per_day$steps))
```

7. Histogram of the total number of steps taken each day after missing values are imputed
==============

Repeat from step 4. with updated dataset produced in step 6. 

```{r}
hist(imp_steps_per_day$steps, main = "Total Steps per day", xlab = "Steps", ylim = c(0,40), labels = TRUE)
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
==============

Add a variable wkdy to the dataset to determine whether this is a weekday or a weekend

```{r}
imputed_data$date <- as.Date(imputed_data$date)
imputed_data$wkdy <- "weekday"
imputed_data$wkdy[weekdays(imputed_data$date) == "Saturday" | weekdays(imputed_data$date) == "Sunday"] <- "weekend"
imputed_data$wkdy <- as.factor(imputed_data$wkdy)
imputed_data_interval <- aggregate(steps ~ interval + wkdy, imputed_data, mean, na.rm = TRUE)
```

Create the plot

```{r}
g <- ggplot(imputed_data_interval, aes(interval, steps))
g + facet_grid(wkdy ~ .) + geom_line() + ggtitle("Average number of steps per 5-min interval")
```
